---
output: github_document
bibliography: biblio.bib
    
nocite: | 
  @Brenning2012, @Russ2010a, @Brenning2005, @Russ2010b
  
csl: apa-old-doi-prefix.csl
---
```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/sperrorest)](http://cran.r-project.org/package=sperrorest)  
[![Build Status](https://travis-ci.org/pat-s/sperrorest.svg?branch=master)](https://travis-ci.org/pat-s/sperrorest) (master)    
[![Build Status](https://travis-ci.org/pat-s/sperrorest.svg?branch=dev)](https://travis-ci.org/pat-s/sperrorest) (dev)     
[![Downloads](http://cranlogs.r-pkg.org/badges/sperrorest?color=brightgreen)](http://www.r-pkg.org/pkg/sperrorest)

# sperrorest

Spatial Error Estimation and Variable Importance

This package implements spatial error estimation and permutation-based
spatial variable importance using different spatial cross-validation
and spatial block bootstrap methods. To cite `sperrorest` in publications,
reference the paper by @Brenning2012.

## Installation

Get the released version from CRAN:
```{r, eval = FALSE}
install.packages("sperrorest")
```

Or the development version from Github:

```{r, eval = FALSE}
devtools::install_github("pat-s/sperrorest", build_vignettes = TRUE)
```


# References




