---
output: github_document
bibliography: "vignettes/Biblio.bib"
biblio-style: apalike
    
nocite: | 
  @Brenning2012, @Russ2010a, @Brenning2005, @Russ2010b
  
---
```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
# description <- readLines("DESCRIPTION")
version <- "1.0.0.9000"
```

```{r, echo = FALSE}
# description <- readLines("DESCRIPTION")
rvers <- "2.10"
# rvers <- stringr::str_match(grep("R \\(", description, value = TRUE), "[0-9]{1,4}\\.[0-9]{1,4}\\.[0-9]{1,4}")[1,1]
```

#### General
[![Build Status](https://travis-ci.org/pat-s/sperrorest.svg?branch=master)](https://travis-ci.org/pat-s/sperrorest)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/pat-s/sperrorest?branch=master&svg=true)](https://ci.appveyor.com/project/pat-s/sperrorest)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![codecov](https://codecov.io/gh/pat-s/sperrorest/branch/master/graph/badge.svg)](https://codecov.io/gh/pat-s/sperrorest)
[![minimal R version](https://img.shields.io/badge/R%3E%3D-`r rvers`-6666ff.svg)](https://cran.r-project.org/)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-yellowgreen.svg)](/commits/master)

#### CRAN
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/sperrorest)](https://cran.r-project.org/package=sperrorest)
[![Downloads](https://cranlogs.r-pkg.org/badges/sperrorest?color=brightgreen)](https://www.r-pkg.org/pkg/sperrorest)
![](https://cranlogs.r-pkg.org/badges/grand-total/sperrorest)

#### Github
[![packageversion](https://img.shields.io/badge/Package%20version-`r version`-orange.svg?style=flat-square)](commits/master)

# sperrorest

Spatial Error Estimation and Variable Importance

This package implements spatial error estimation and permutation-based
spatial variable importance using different spatial cross-validation
and spatial block bootstrap methods. To cite `sperrorest` in publications,
reference the paper by @Brenning2012.

## Installation

Get the released version from CRAN:
```{r, eval = FALSE}
install.packages("sperrorest")
```

Or the development version from Github:

```{r, eval = FALSE}
devtools::install_github("pat-s/sperrorest", build_vignettes = TRUE)
```

# References




